---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(Lahman)
library(mosaic)
library(tidyr)
library(tidyverse)
library(dplyr)
```

```{r}
#Load in People, Batting, and Pitching Dataframes
data("People") 
data("Batting")
data("Pitching")
```

```{r}
#Merges player name to Batting data. 

People$name <- paste(People$nameFirst, People$nameLast, sep = " ")

batting <- merge(Batting,
                 People[,c("playerID", "name")],
                 by = "playerID", all.x = TRUE)

#Merges player name to Pitching data.

People$name <- paste(People$nameFirst, People$nameLast, sep = " ")

pitching <- merge(Pitching,
                 People[,c("playerID", "name")],
                 by = "playerID", all.x = TRUE)
```

## Data Preparation (Lesson 1 & 2)

```{r}
#Keep players with over 100 at bats. (We can change this value if necessary).
#Creating batting average variable.

batting <- batting %>%
  filter(AB >= 100) %>%
  mutate(BAVG = (H/AB))
  
```


## Exploratory Analysis (Lesson 1 & 2)
Lessons 1 and 2 will just be parts of the overall project. Simple things like data manipulation, apply functions, boxplots, etc. This will be data preparation items and exploratory analysis.

```{r}
b <- ggplot(batting, aes(x = teamID, y = HR)) +
  geom_boxplot(col = "black", aes(fill = teamID))
b

```

## 0-dimensional Reduction (Lesson 4)

#### Bootstrapping

## PCA (Lesson 4)

#### Screeplot

#### Biplot

## Cluster Analysis (Lesson 5)

#### Silhouette

#### Diana

## Linear Regression (Lesson 6)

## Simple Linear Regression

## Multiple Linear Regression
